---
title: "Lab Notebook: September 2018"
output: html_document
---
#Goals: September 17 - 22
1. Authorship paper:
    + ~~Resubmit job to link/verify manuscript group ids~~
    + Link genderize
    + Finish script to pull author names from URLs
    + Test Nichole's dataset in genderize

1. Monthly journals report: 
    + ~~Ask Melissa to check in with Joel~~
    + ~~Email to ask about encryption on Flux~~

1. ~~BA manuscript - finalize revisions~~

1. ASM blog post
    + ~~draft pathogenic fungi post~~
    + make revisions
    
1. Software carpentries workshop
    + contribute to the lessons
    + work through R for Social Scientist lessons
    + ~~reschedule discussion & demo~~
    
#2018/09/19

**- Lab meeting**

**- M&I newsletter**

Coordinate photo w. Denise & Michele. Edit stuff. 

**- Authorship Paper**

Read paper examining peer review bias at eLife. They use a consultative peer review process where after the reviewers submit write ups, they have a discussion about the merits of the paper (where identies are known) prior to a decision being given. 

Thoughts:
- histogram of # publications per unique id, M vs F vs Unknown
- chi-squared comparision between groups
- total M v F compared to over time, any changes? 2012 vs 2017
- fully explain ASM submission/review process in intro/data
- all M reviews vs all F reviews vs mixed/uncertain
- author/reviewer country origin (& gender) look at reviewer reccomendation/score instead of (or in addition to) editor decision
- no objective measure of the intrinsic quality of a manuscript (citations? narrative?)
- authors don't discuss how the eLife method could perpetuate homophily by potentially allowing one (high ranking) reviewer to dominate the discussion

**- Monthly journals report**

Sent another email to Joel, cc'ing Pat & Melissa.

Successfully encrypted a directory on flux. Now I'm confused about the de-encryption process. I have to enter a passphrase to encrypt a file but none to de-encrypt it? I guess this is as long as I'm in the same terminal session. If I log out of flux, then log back in I have to enter the passphrase.

Need to set options to use a better encryption method: https://superuser.com/questions/633715/how-do-i-fix-warning-message-was-not-integrity-protected-when-using-gpg-symme

Need to work on building the script to de/re encrypt the files for analysis.

More reading on security/encryption/keypairs:
- https://access.redhat.com/documentation/en-us/red_hat_enterprise_linux/6/html/security_guide/chap-security_guide-security_overview
- https://www.digitalocean.com/community/tutorials/how-to-use-gpg-to-encrypt-and-sign-messages
- https://alexcabal.com/creating-the-perfect-gpg-keypair/
- https://www.gnupg.org/gph/en/manual.html

**- BA manuscript**

Finalized changes to the manuscript, uploaded changes plus reviewer comments to mBio site. Just waiting for the cover letter.

#2018/09/18

**- Lab inventory**

**- Software carpentries**

Rescheduled discussion (next week, need to email host) & demo.

**- Monthly journals report** X

Work on encrypting/de-encrypting files. I was able to encrypt/de-encrypt a txt doc on flux with help from this link: https://www.if-not-true-then-false.com/2010/linux-encrypt-files-decrypt-files-gpg-interactive-non-interactive/

```{r}
gpg -c filename.txt #to encrypt

gpg filename.txt.gpg #to de-encrypt
```

Unfortunately, dealing with entire directories is more complex. I have to first compress the directory in to a `tar.gz` file that can then be encrypted. https://statistics.berkeley.edu/computing/encrypt

```{r}
tar czf my_dir.tar.gz my_dir/ #compress to tarball that can be encrypted
  
tar xzf my_dir.tar.gz #unzip back to directory
```

**- M&I newsletter** XX

Made suggested edits to the docs currently available. 

**- ASM blog post** XXXX

Finish & submit.

#2018/09/17

**- Lab Meeting**

Machine Learning. 

**- ASM blog post**

Work on draft.

**- Authorship Paper**

Check fix_link job submitted. It ran the first part, but then ran out of wall time to run the testing portion which means it didn't save the corrected csv. Moved up saving the file in the csv, added walltime & resubmitted the job.

**- M&I Newsletter**

Edits to Heritage story

**- EJP access**

Tested encrypting files in flux, permission denied. Need to check in w. admin (command worked locally)

**- Prof Development**

Consider applying to DEI certificate program. Application due Friday:

"Please explain why you are interested in completing the Professional Development DEI Certificate Program, and how you believe that the certificate will help you to achieve your professional development goals. Please limit your response to 500 words."

#Goals: September 10 - 14
1. Authorship paper:
    + ~~Link manuscripts~~ 
    + Link genderize
    + Finish script to pull author names from URLs
    + Test Nichole's dataset in genderize

1. Monthly journals report: 
    + Ask Melissa to check in with Joel

1. BA manuscript - finalize revisions (checked in w. Phil)

1. ASM blog post
    + draft pathogenic fungi post
    + ~~suggestions to this week's post~~
    
1. Software carpentries workshop
    + contribute to the lessons
    + work through R for Social Scientist lessons

1. ~~M&I Newsletter - Society News (ASM/ASV)~~
    
1. ~~Help @ SWC - Thursday 9 - 4:30~~

#2018/09/13

**- Help @ SWC workshop**

**- ASM blogging**

Work on draft

#2018/09/12

**- Lab meeting**

**- M&I newsletter**

Suggested edits to stories currently available

Wrote draft of Society News

**- ASM blog post**

Work on draft

#2018/09/11

**- ASM blogging**

Compiled list of permission requirements by publisher.

**- SWC** X

Look through lesson plans & current issues.
lessons - https://datacarpentry.org/r-socialsci/
instructor notes- https://datacarpentry.org/r-socialsci/guide/

**- ASM blog post**

Work on outline for pathogenic fungi. XXXX

Begin draft and rearrange notes into a slightly better outline. XX

**- M&I Newsletter** XX

Edits/stories due by 10/3 - Due to printer 10/22

Made suggested edits to a couple of stories emailed by Kathy (DEI & Zambia).

Emailed Denise about coordinating a photo for the newsletter. Started drafting society story.

**- Fall/Summer Strategic Goals** XX

Reworked rest of Summer goals & started Fall goals

#2018/09/10

**- Lab meeting**

**- Misc**

Worked on revamping summer strategic goals.

**- ASM blog post** XXX

Make suggestions to Rita's gliobastoma post.

**- Authorship Paper** XXXX XX

Write script to fix mis-matched grouped manuscript numbers & verify that none remain. Submit to flux. Script required generating a df of all manuscripts with the wrong group number in one column & the correct grouped number in a second column. This df was joined to the grouped dataframe (`all_grouped`), then the `grouped.manu.number` corrected as necessary. Ran a second verification to confirm that there aren't any matches remaining.

#Goals: September 3 - 7
1. Authorship paper:
    + Draft methods
    + Link manuscripts 
    + Finish script to pull author names from URLs
    + Test Nichole's dataset in genderize
    + ~~Pull datasets from GitHub~~

1. Monthly journals report: 
    + Report mockup w. AEM data
    + ~~Check in with Joel about the data transfer~~

1. ~~BA manuscript - finish revisions~~

1. ASM blog post
    + draft pathogenic fungi post
    + ~~suggestions/edits for Monika~~
    
1. Software carpentries workshop
    + ~~schedule final assignments~~
    + contribute to the lessons
    
1. M&I department picnic - Friday, 1 - 5pm

1. ABT workshop for 812 

1. M&I Newsletter
    + ~~Meet w. Kathy - Monday @ 10:45~~
    + ~~Write Neidhardt-Freter 2017 story~~
    + Society News (ASM/ASV)
    
1. ~~Irene's Thesis Defense - Thursday @1pm, NLH~~

#2018/09/07

**- M&I Dept Picnic**

**- ABT workshop w. 812**

**- Authorship Paper** XX

Script I submitted yesterday to check the xml matches failed b/c I screwed up `write_csv`. Fixed it and resubmitted the batch script.

Worked on script to webscrape author names. Figured out a trick for identifying journal from pubmed, but I think I'm going to have to add a line earlier in the script to specify journal, then use that to call functions later. 

Remaining journals: science, wiley library, other misc.

Interactive job w. Rstudio/Flux: https://arc-ts.umich.edu/arc-connect/

#2018/09/06

**- Practice ABT talk with Nick**

**- Irene's Thesis Defense**

**- Authorship paper** XXX

Pull `all_grouped.csv` from flux. Trying to figure out which manuscripts are getting multiple manuscript numbers attached by writing a function to count the number of identifiers a number matches to. I can then select for those with >1 match & look at them individually. There should be less than 50 so it shouldn't be an issue.

Next step is to pull the manuscript number & grouped manu number columns, then join to other datasets (people & reviews) so that they can be tracked similarly.

After that, check that gender data is merging properly & finally... analyze data!!

Looked at genderized names, picked a threshold of 65%.

#2018/09/05

**- Lab meeting**

R hacks: data-to-vis.com, r-graph-gallery.com, 

**- SCW**

Signed up to teach in October & signed up for discussion & teaching demo.

**- Authorship Paper**

Fixed `link_manu` script & re-submitted batch script.

**- ASM blog**

Work on outline for pathogenic fungi post

#2018/09/04

**- ASM blog post**

Made suggestions/edits to Monkia's Zika post.

**- M&I Newsletter**

Met with Kathy & wrote a draft of the 2017 NF symposium story.

**- Lab Stuff**

Moved desks & went to get new keys

**- Authorship paper**

Worked on webscraping script & sent to Pat for suggestions.

**- Monthly journals report**

Figured out how to intiate `sftp` using flux. Created back ups of JIf, Impact Vizor & Reports folders since I'm going to delete those GitHub repositories. 

```
> sftp akhagan@flux-xfer.arc-ts.umich.edu:/nfs/turbo/schloss-lab/akhagan/
#enter password

sftp> put -r Desktop/estimated_jif #transfer a local file to remote server

sftp> get -r remoteFile localFile #or pull a file from the remote server and put locally
```

#2018/09/03

**- BA manuscript**

Finished revisions & sent to Phil for comment